---
title: Academic Activities & CV
toc: true
---

```{r, include = FALSE}
source("pre-render.R")
```


> I was honoured to receive the Society of Canadian Ornithologists/Société des ornithologistes du Canada's inaugural **[Early Career Research Award](https://www.sco-soc.ca/ecr)** for 2017


## Academic Activities 

#### Positions
**2018-Present** Adjunct Professor, Department of Biology, Brandon University

#### MSc Committees
**2019-Present** Department of Geography & Environment, Brandon University  
**2019-Present** Department of Biology, Brandon University  
**2019-2020** Department of Biological Sciences, University of Manitoba  


## Education

__PhD (2010 – 2015)__
University of Northern BC, Natural Resources and Environmental Studies  
Supervisor: Dr. Ken A. Otter  
Thesis: [Chickadee vocalizations in human-altered environments](http://www.unbc.ca/sites/default/files/sections/ken-otter/lazerte-dissertation2015-04-28final.pdf)

__MSc (2007 – 2010)__
McGill University, Biology  
Supervisor: Dr. Don L. Kramer  
Thesis: [Using thermosensitive telemetry to measure activity in eastern chipmunks (_Tamias striatus_)](http://digitool.library.mcgill.ca/R/-?func=dbin-jump-full&object_id=95000&silo_library=GEN01)

__BSc (2003 – 2007)__
University of Toronto, Honours Zoology, Behaviour Specialist  
Supervisor: Dr. Allan Baker  
Independent Project: Genomic scan using AFLPs reveals that red knot subspecies are genetically differentiated

## Post-Docs

__Post-Doc (2016 – 2017)__
Thompson Rivers University, Geography and Environmental Studies  
Supervisor: Dr. David J. Hill  
Details: Developing citizen science tools, in particular an open system for the web-based visualization and analysis of animal movement data

__Post-Doc (2015 – 2016)__
University of Northern BC, Natural Resources and Environmental Studies  
Supervisor: Dr. Ken A. Otter  
Details: Tracking gap-crossing in small birds through automated RFID feeders.

__Post-Doc (2015)__
University of Northern BC, Natural Resources and Environmental Studies  
Supervisor: Dr. Ken A. Otter  
Details: A follow up study to playback studies conducted during my PhD looking at audibility of different vocalizations in noisy environments.

## Publications

<script src="https://bibbase.org/show?bib=https%3A%2F%2Fapi.zotero.org%2Fusers%2F1841%2Fcollections%2FFNZPV2RA%2Fitems%3Fkey%3DIuq7CUIMuNjpsrwtNRycDTsQ%26format%3Dbibtex%26limit%3D100&group0=year&nocache=1&authorFirst=1&jsonp=1"></script>


## Presentations

<script src="https://bibbase.org/show?bib=https%3A%2F%2Fapi.zotero.org%2Fusers%2F1841%2Fcollections%2FP9S6VACZ%2Fitems%3Fkey%3DIuq7CUIMuNjpsrwtNRycDTsQ%26format%3Dbibtex%26limit%3D100&group0=year&nocache=1&authorFirst=1&jsonp=1"></script>


## Teaching

### R University Teaching

```{r}
extra %>%
  filter(type == "r teaching") %>%
  fmt_teaching(type = "html") %>%
  fmt_gt()
```

### R Workshops
```{r}
workshops %>%
  mutate(title = str_remove(title, " Workshop")) %>%
  fmt_teaching(type = "html") %>%
  fmt_gt()
```

### R Volunteering
```{r}
tibble(start = c(2012, 2012), 
       end = c("Present", 2017),
       content = c(
         glue("**One-on-one assistance**\n",
              "- University of Manitoba, Winnipeg, MB\n",
              "- Brandon University, Brandon, MB\n",
              "- University of New Brunswick, Fredericton, NB\n",
              "- Wilfrid Laurier, Waterloo, ON\n",
              "- University of Northern BC, Prince George, BC"),
         glue("**Organized R Help Groups**\n",
              "- R Group (2016-2017) {small('University of New Brunswick, Fredericton, NB', 'html')}\n",
              "- R Coffee Hour (2012-2016) {small('University of Northern BC, Prince George, BC', 'html')}"))) %>%
  fmt_gt() |>
    cols_width(1 ~ "10%", 
               2 ~ "10%")
```

### Other University Teaching
```{r}
extra %>%
  filter(type == "teaching") %>%
  fmt_teaching(type = "html") %>%
  fmt_gt()
```


#### Large events
- 2015 __UNBC Discovery Centre__  
This centre was a 3-day event held in November 2015. I was one of three coordinators and organized events, volunteers and panelists, and supervised on all event days.


#### Attendance at Teaching Conferences and Workshops
```{r}
tribble(
  ~start,     ~title, ~location,
  "2018-08", "Brandon University Teaching Excellence Conference", "Brandon, MB", 
  "2011–2016", "UNBC and CNC Annual Teaching and Learning Conference", "Prince George, BC",
  "2014-09", "Teaching Dossier Workshop", "Prince George, BC", 
  "2011-08", "Instructional Skills Workshop", "Prince George, BC",  
  "2011-01", "UNBC Teaching Assistant Workshop", "Prince George, BC",   
  "2008-03", "McGill Professional development workshop: Learning to Teach", "Montreal, QC", 
  "2008-01", "Teaching Workshop – McGill Tomlinson Project in University-Level Science Education", "Montreal, QC") %>%
   fmt_other(type = "html") %>%
  fmt_gt()
```


#### Community presentations and outreach
 
```{r}
tribble(
  ~date, ~title, ~role, ~desc, ~location,
  "2015-01", "Prince George Naturalists Club", "Guest speaker", "Singing through the noise: How birds compensate for effects urbanization on communication", "Prince George, BC",
  "2013-11", "Williams Lake Naturalists Club", "Guest speaker", "Singing through the noise: How birds compensate for effects urbanization on communication", "Williams Lake, BC",
  "2014-05", "Prince George Young Naturalists Club", "Guest speaker", "Come learn all about Chickadees!", "Prince George, BC",
  "2014-03", "Immaculate Conception School Science Fair", "Judge", NA, "Prince George, BC",
  "2013-11", "Lake City Secondary School Science 8 and 9", "Guest Educator", "Early bird gets the worm!", "Williams Lake, BC", 
  "2013-10", "UNBC Northern Research Group Trail Talks", "Guest Guide", "Birds and Glacial Lakes", "Prince George, BC",
  "2013-09", "Quesnel River Research Centre High School Salmon Day", "Guest Scientist", "Chickadee communication and sound transmission activities",  "Likely, BC",
  "2012/09 & 2013/09", "Kelowna Kokanee Salmon Festival", "Exhibitor", NA, "Kelowna, BC") %>%
  fmt_other(type = "html") %>%
  fmt_gt()
```
 

## Awards and Scholarships

- SCO-SOC [Early Career Researcher Award](https://www.sco-soc.ca/_files/ugd/c932da_1f8708c113e74becb953548b95b13020.pdf) (2017/08)
- NSERC Postgraduate Scholarship (PGS) Doctoral (2011/05 – 2014/05)
- Society of Canadian Ornithologists (SCO-SOC) First Place Student Presentation (2013/08)
- Society of Canadian Ornithologists (SCO-SOC) James L. Baillie Award (2012)
- NSERC Alexander Graham Bell Canada Graduate Scholarship (CGS) Masters (2008/04 – 2009/04)
- NSERC Undergraduate Student Research Award (2006)

## Media

Over the years several projects I've been involved with have been fortunate enough to attract some media attention.

### Chickadees and Noise

- __2016/07__ Media attention surrounding publication of "Learning to cope: Vocal adjustment to urban noise is correlated with prior experience in black-capped chickadees"  
  - TV: [CKPG (watch)](http://www.ckpg.com/2016/07/11/unbc-scientist-makes-new-discovery/)
  - Radio: [CBC Daybreak North (listen)](images/2016-07-12 - CBC Daybreak.mp3)
  - Newspaper: [Prince George Citizen (read)](http://www.princegeorgecitizen.com/news/local-news/unbc-researcher-shows-chickadees-adapt-song-to-city-1.2299671)
  - Newsite: [250news.org (read)](https://www.250news.com/2016/07/11/study-shows-urban-rural-chickadees-chirp-to-a-different-tune/)

- __2013/09__ SCO award for best student presentation highlighted by Ducks Unlimited
-- [Ducks Unlimited facebook post](https://www.facebook.com/ducksunlimitedcanada/posts/10151864386454800)

- __2012/04__ Sharing relevant findings with the Stanley Park Ecological Society -- [Stanley Park Ecological Society (read)](http://stanleyparkecology.ca/2012/04/09/vancouvers-chickadees-sing-a-different-tune-2/)


### White-throated Sparrows Citizen Science

- __2016/03__ Start of the citizen science study -- [CKPG (watch)](http://www.ckpg.com/2016/03/28/sparrows-singing-a-new-tune/)
- __2016/05__ Geolocators and tracking movements -- [CKPG (watch)](http://www.ckpg.com/2016/05/26/live-capture-and-release/)

