---
title: "Running Basic Linear Models"
date: "2022-11-13"
categories: [stats, lm, R]
---

In R, basic linear models are run with the `lm()` function. 

```{r, eval = FALSE}
lm(y ~ x1 + x2, data = data)
```

R will choose which type of linear model to do depending on the type of values 
in `x1` and `x2`.

- **Regression** - One continuous predictor (`x1` only and is continuous)
- **Multiple regression** - More than one continuous predictor 
  (both `x1` and `x2` are continuous)
- **ANOVA** - Categorical predictor(s) (both `x1` and `x2` are categorical - factors)
- **ANCOVA** - Mix of categorical and continuous predictors (`x1` is continuous and `x2` is categorical)


In all cases, `y` is continuous.  

No count data, no binary (binomial) data, no proprotions.

[(Although technically sometimes count data is close enough to continuous and 
proportions can be transformed)]{.small}


### Quick example

```{r}
library(palmerpenguins) # To access the penguins data set
m <- lm(body_mass_g ~ bill_depth_mm, data = penguins)
m
```

> But wait! How do I interpret that?

All the details are stored in the model object `m`. We can use many other
R functions to check our assumptions, summarize the results, or create ANOVA tables.

But first, you really should check your [Diagnostics](../diagnostics-lm)!

### Side note on `NA`'s

By default missing values (`NA`'s) are dropped from the analysis and from the
data contained by the model. However, even better is to use `na.exclude`.

```{r}
m <- lm(body_mass_g ~ bill_depth_mm, data = penguins, na.action = na.exclude)
```

This also excludes missing values (`NA`'s) from the analysis, but *not* from the
data you extract from the model (which makes it much easier to combine back with
your raw data set when performing [Diagnostics](../diagnostics-lm)).


### Session Info
```{r}
devtools::session_info()
```

